---
title: "HW 9 - Anaya Malik"
output: html_document
date: "ahm2466"
---

[GitHub Link Here!](https://github.com/anayaM800/HW9.git)


### Question 1
```{r setup, include=FALSE, echo=FALSE}
library(dplyr)
library(ggplot2)
library(tidyverse)
library(mosaic)
library(MatchIt)


# Part A
turnout <- read.csv('turnout.csv')

GOTV_1998 <- select(turnout, GOTV_call, voted1998)
prop_received_call <- mean(GOTV_1998[GOTV_1998$GOTV_call == 1, "voted1998"])
prop_no_received_call <- mean(GOTV_1998[GOTV_1998$GOTV_call == 0, "voted1998"])
prop_received_call
prop_no_received_call


xtabs(~GOTV_call + voted1998, data=GOTV_1998) %>%
  prop.table(margin=2)

prop.test(GOTV_call ~ voted1998, data=turnout, success = 1)



```
#### PART A
The prop of those getting a call was 0.6477733 and the propl of those not getting one was 0.4442449. The conf. interval being -0.024 and -0.012 between those who didnt get a call and those who did. 


```{r, echo=FALSE}
# Part B
xtabs(~GOTV_call + voted1998, data=turnout) %>%
  prop.table(margin=2)


prop.test(GOTV_call ~ voted1998, data=turnout)


mean(AGE ~ GOTV_call + voted1998, data=turnout)
mean(voted1996~ GOTV_call + voted1998, data=turnout)
mean(MAJORPTY ~ GOTV_call + voted1998, data=turnout)

ggplot(turnout, aes(x = AGE, fill = factor(GOTV_call))) +
  geom_histogram(position = "dodge", bins = 30) +
  facet_wrap(~voted1998) +
  labs(title = "Age Distribution by GOTV Call and Voting Status", x = "Age", y = "Count")

ggplot(turnout, aes(x = factor(voted1996), fill = factor(GOTV_call))) +
  geom_bar(position = "dodge") +
  facet_wrap(~voted1998) +
  labs(title = "1996 Voting by GOTV Call and Voting Status", x = "Voted in 1996", y = "Count")

ggplot(turnout, aes(x = MAJORPTY, fill = factor(GOTV_call))) +
  geom_bar(position = "dodge") +
  facet_wrap(~voted1998) +
  labs(title = "Major Party Affiliation by GOTV Call and Voting Status", x = "Major Party", y = "Count")


```


#### PART B
To show that these three factors are confounders, firstly you can see the assoication of each of the factors through the graph. The younger crowd for exmaple tended to vote more when not called while the older crowd tended to vote more when called. Between those who voted in 1996 there is also association as for example those who did vote in 1996 increasingly voted in 1998 if they did get a call. However if they didn't vote in 1996, they seemed to vote less after a call. Similarly for majority party, the trend for both, if they are in majority part or not, a call seemed to increase the voters for 1998. From the outputed test, the x squared, df and p value show that there is statistical signifance in the difference of people who voted in 1998 from those who got a call and didn't but to say its because of one of those factors might be diffucult to say. Althought influential, they may not be causation. The conf interval is also stat sig, again saying that getting a call can increase voting but those factors remain cofounders as they are not causasional.


```{r pressure, echo=FALSE}
# Part C

matched = matchit(GOTV_call ~ voted1996 + AGE + MAJORPTY,
                                 data = turnout, ratio=5) %>%
  match.data

mean(AGE ~ GOTV_call + voted1998, data=matched)
mean(voted1996 ~ GOTV_call + voted1998, data=matched)
mean(MAJORPTY ~ GOTV_call + voted1998, data=matched)


match_GOTV <- select(matched, GOTV_call, voted1998)
prop_received_call_matched <- mean(match_GOTV[match_GOTV$GOTV_call == 1, "voted1998"])
prop_no_received_call_matched <- mean(match_GOTV[match_GOTV$GOTV_call == 0, "voted1998"])
prop_received_call_matched
prop_no_received_call_matched

prop(voted1998~ GOTV_call, data=matched, success = 1)
prop.test(voted1998 ~ GOTV_call, data=matched, success = 1)



```

#### PART C
With the confounders balanced and similarity across the numbers this ensures balancing. The proportion of those receiving a GOTV call who voted in 1998 is 0.6477733 and the proportion of those not receiving a GOTV call who voted in 1998 is 0.5740891. This means that those who got a call are more likley to vote than those who didn't. The stats show the conf. interval being  -0.14 and -0.0056 showing a negative difference meaning that those who got calls voted more than those who did not. After balancing it out, the call seems to positivley impact the voting rates in 1998. 


### Question 2
```{r, echo=FALSE}
solder <- read.csv('solder.csv')

# Part A

ggplot(solder, aes(x = Opening, y = skips)) +
  geom_boxplot() +
  labs(title = "Relationship Between Solder Opening Size and Number of Skips",
       x = "Opening Size",
       y = "Number of Skips")


ggplot(solder, aes(x = Solder, y = skips)) +
  geom_boxplot() +
  labs(title = "Relationship Between Solder Thickness and Number of Skips",
       x = "Solder Thickness",
       y = "Number of Skips")

```

#### PART A
In the first plot, the size of the opening is related to the number of skips as you can see as the size becomes progressivley smaller the numbers of skips data becomes more spread. The smaller it is, it seems as if the number of skips is more frequent. The average number of skips also increases with each size going from largest to smallest. 

In the second plot, the thickness is ploted along with the number of skips and they are related in the sense that the thin solder has a higher average and more variability for the number of skips as the higher number of skips seems to be related to thinner solder. With thick, the average number of skips is lower with the thicker ones and the higher number of skips of data for thicker ones seems to be less than the higher points for the thinner ones. 

```{r, echo=FALSE}
model <- lm(skips ~ Opening + Solder + Opening:Solder, data = solder)
coef <- coef(model)
interval <- confint(model)
table <- cbind(coef, interval)
colnames(table) <- c('Estimate', '2.5%', '97.5%')
knitr::kable(table)

```


#### PART B & C
Above, this table was created with the help of a linear regression model and 95% confidence for each coeff. Each intercept has the estimate, 2,5 and 97.5 listed. This table was created for part b of the question. For part C, to interpret each of the estimates, the first estimate for the intercept if 0.39 when all other predictors are 0. Using an opening of M, the estimate skips is 2.41 while small is 5.12. The thin over the thick ones are 2.28 skips and the thin opening for medium is -0.74 skips and for small thin its 9.65. 


#### PART D
For part D, if there had to be one choice, I would chose the thinner medium because the estimate value is the best for this combination compared to other types. It does not render statistical signifcance because the conf. interval does indeed include zero, but practically speaking the estimate numbers are the best. If you did want to value statistical evidence though and not include a zero in the conf. interval then you could chose a slighter bigger estimate with a conf. that dosn't include zero. Another choice that could make sense is the intercept as it has a small skip rate also. Even though the confidence interval is not stat significant, this one has a smaller confidence interval meaning greater certainity. This is one is also easier to see with the proof of graphs for low skip rate. 







